Effects of Tunable Hydrophobicity on the Collective Hydrodynamics of Janus Particles under Flows

### 2022-08-05 doc/PRF2022/PRF_main.tex
yny

%Title of paper
\title{Collective behavior of Janus Particles Suspended in a Viscous Fluid}

@@

\begin{abstract}
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in biology. To understand how
  hydrodynamic flow may be used to actively control self-assembly of
  Janus particles (JPs), we use a model recently developed for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  background flow (JFM, 941, 2022). We investigate how various
  morphologies arise from tuning the hydrophobic distribution of the
  JP-solvent interface. We find JPs assembled into uni-lamella,
  multi-lamella and striated structures. To introduce dynamics, we
  include a linear shear flow and a Taylor-Green mixing flow, and
  measure the collective dynamics of JP particles in terms of their (a)
  free energy from the hydrophobic interactions between the JPs, (b)
  order parameter for the ordering of JPs in terms of alignment of their
  directors, and (c) strain parameter that captures the deformation in
  the assembly. We characterize the effective material properties of the
  JP structures and find that the uni-lamellar structures increases
  orientation order under shear flow, the multilamellar structure
  behaves as a shear thinning fluid, and the striated structure
  possesses a yield stress. These numerical results provide insights
  into dynamic control of non-equilibrium active biological systems with
  similar self-organization.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations: Hydrophobic Attraction Potential Mobility Problem} 
The governing equations are a nonlinear system for the position and
orientation of a collection of rigid Janus particles. We first pose the
Stokes equations for the mobility problem giving the hydrodynamic
interactions for the particle suspension. The hydrophobic forces come
from solving a screened Laplace equation. Particle collisions are
avoided through a near-field, pair potential.

@@

To perform a single time step, we first solve for the HAP by
solving~\eqref{eq:SL}--\eqref{eq:SLbc}, and then compute the force and
torque~\eqref{eq:FandTdef}. Then the Stokes
equations~\eqref{eq:stokes}--\eqref{eq:stokesFarfield} are solved, and
the force-free and torque-free conditions are used to compute the rigid
body motion~\eqref{eq:rigid-vel}.
Once~\eqref{eq:stokes}--\eqref{eq:force} have been solved, the
velocities $(\vv_i, \omega_i)$ from~\eqref{eq:rigid-vel} update the
particle positions and orientations.

@@

\section{Model parameters and boundary conditions} The interactions
generated by the HAP-mobility problem formulation
\eqref{eq:stokes}--\eqref{eq:force} lead to particle self-assembly for a
broad range of parameters and boundary conditions. Generally speaking,
the effective distance of the hydrophobic interaction is set by the
screening length $\rho$. Smaller values of the repulsion distance
$\rho_0$ decrease the distance where attraction and repulsion are in
balance and lead to more compact particle assemblies. The rate of
self-assembly is proportional to the interfacial tension $\gamma$, and
inversely proportional to solvent viscosity $\mu$; it is roughly
inversely proportional to screening length $\rho.$

@@

We set $\max g>0$ to guarantee at least partial hydrophobicity of the JP
surface. Then, the boundary condition (BC) can be classified into three
categories: $\min g(\xx) = 0$, $\min g(\xx) >0$, and $\min g(\xx) <0$
where the extrema are taken for $\xx \in \Gamma_i$. These three
categories correspond to three characteristic values of $b$ in
\eqref{eq:bc-type}: $b=1$, $b=2$ and $b=0$, respectively. 
%
To see the effect of the shift parameter $b$, we simulate the dynamics
of 198 particles and 60 particles in a quiescent flow as shown in
Figure~\ref{fig:relax}. Three distinct configurations emerge: (i)
bilayer for $b = 1$, (ii) multilamellar for $b = 2$, and (iii) striated
for $b = 0$. Supplementary Movie S1 shows the self-assembly process for
the three cases.

@@

Figure~\ref{fig:relax}(a) shows the initial configurations with
orientations of the 198-particle and 60-particle cases, respectively.
BC (i) with $b = 1$ mimics an amphiphilic particle. The boundary data
$g$ is everywhere nonnegative, and takes the maximum value $a$ on the
hydrophobic tail (red) the value $0$ on the hydrophilic head (green).
The interaction is attractive, and particles collectively orient their
hydrophobic sides to form bilayers (Figure~\ref{fig:relax}(b)). The
inset shows the equilibrium state for the 60-particle run. Both
$N_b = 60$ and $N_b = 198$ cases
form bilayers with multiple components.

@@

\section{Measuring deformation}
We use the free energy $F$, a strain parameter $E$, and two scalar order
parameters $S_{2}$ and $\tilde S_2$ to measure the deformation of
particle configurations under background flows.

@@

Finally, we use the scalar order parameter for orientational order\cite{Selinger2016}:
\begin{equation}
  \label{eq:S2}
S_2 = \frac{1}{N_b} \sum_{i=1}^{N_b} \frac{1}{2}(3\cos^2(\theta_i - \bar \theta) - 1).
\end{equation}
Here $\bar \theta$ is a circular mean defined as the orientation of the
principal eigenvector of the matrix $\dd_1\dd_1^\top + \cdots +
\dd_{N_b}\dd_{N_b}^\top$. A value $S_2(t) = 1$ indicates that all
particle directors lie on a common axis e.g., two parallel vectors with
possibly opposite direction are ordered. A value far from 1 indicates
disorder.

@@

%%%%%%%
%
\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{Figures/Figure5.pdf}    
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:BC1_TG} Multiple component bilayers in a
  Taylor-Green when $V_0 = \{0.1, 0.15, 0.2\}$ at $t=0.2\ \upmu$s. The
  pre-relaxed initial configuration is shown in the inset of panel (a).
  The streamlines are plotted in the background. Panel (d) shows the
  free energies; panel (e) shows the orientational parameter
  $\tilde{S}_2$; panel (f) shows the positional parameter $E$.}
\end{figure}
%
\subsection{Bilayer and vesicle configurations in a Taylor-Green flow}
Next, we subject the JP structures to the Taylor-Green (TG) background
flow
\begin{equation}
\label{eq:tg_flow}
\uu_\infty(\xx) = V_0\; \frac{\text{nm}}{\text{ns}}\;
\left(
-\cos\left(x/\lambda\right)
 \sin\left(y/\lambda\right)
         {\bf e}_x
         +\sin\left(x/\lambda\right)
         \cos\left(y/\lambda\right)
             {\bf e}_y\right),
\end{equation}
where $x = {\bf e}_x \cdot \xx$ and $y = {\bf e}_y \cdot \xx$
are the horizontal and vertical coordinates, respectively. 
The TG flow is a confining flow consisting of a checkerboard pattern
of cells with alternating circulation. We control the flow by the
dimensionless flow strength $V_0$ and the dimensionless cell size $l$
defining $\lambda = l$ nm. For most of the results in this subsection,
we use the parameter $l=2$.
%

@@

There is no appreciable pattern in the deformation for the
multicomponent bilayer case in TG flow
(Supplementary Movie S3, right panel).
Figure~\ref{fig:BC1_TG}(a)--(c)
are the snapshots when $V_0=0.1,0.15,0.2$ at $t = 0.2$ $\upmu$s. At all
flow strengths, the bilayer components are separated and mixed by the
flow in neighboring cells. In the $V_0 = 0.2$ case, the flow breaks off
components consisting of only a few particles, leading an increase in
energy (Figure~\ref{fig:BC1_TG}(d), long dashed curve). There is no
apparente structure to the change in the orientational order. There is an initial
increase in the strain parameter due to component mixing, but the
strains remains bounded since the particles stay confined to the
neighbors of the central cell.

@@

\citet{kohl-cor-che-vee22} have simulated JP suspensions in three
dimensions using the same free energy formulation of hydrophobic
attraction used in the present work. They observed spontaneous
aggregation into micelles generically and formation of bilayers when the
initial configuration was chosen sufficiently close to the final
configuration. They also studied self-assembly of bipolar electric JP,
which use BC (ii) but electric potential, instead of a water order parameter,
to define the interaction.  Due to the differing interaction,
bipolar electric particles form parallel chains that
repel. In contrast, the hydrophobic interaction between stria is
attractive. The existence of 'X'-shaped local arrangements as a local
equilibrium (Figure~\ref{fig:relax}(d), top, middle rectangle) further
suggests there may be greater diversity in the set of possible
equilibria in three-dimensions when employing hydrophobic attraction.

@@

Under a relatively weak linear shear flow, the amphiphilic JPs behave as
a unilamellar vesicle that elongates and tank-treads, with the scalar
order parameter increasing over time. The assembly of multi-lamella and
striated JP structures resemble a rigid body motion with minimal
deformation. The effective viscosity of the material quantitatively
validates this result. High shear-rate cases provide a range of critical
shear rates where the structures break apart and undergo topological
changes. We further show that free energy, scalar order parameter, and
strain are effective measures to quantitatively capture the dynamics of
the JP oligomers under a linear shear flow.

Under a Taylor-Green flow, the amphiphilic JPs are the most interesting,
exhibiting different vesicle shapes depending on the ratio of the size
of the TG flow to the vesicle size. These results show that the shape of
the vesicle, whether square or polygonal, can be controlled by adjusting
the size of the cell in the TG flow. On the other hand, the assembly of
multi-lamella JPs and the striated JPs behave more like a rigid body
with connected subdomains, and the number of subdomains increases with
increasing strength of the TG flow. Overall, the multilamellar (BC (ii))
JPs oligomers behave as a shear-thinning fluid, while the striated (BC
(iii)) JPs oligomers possess a yield stress.

@@

%%%
\subsection{Hydrophobic Attraction Potential Mobility Problem} 
The mathematical formulation is a nonlinear system for the dynamics of a
collection of rigid Janus particles. For the sake of completeness, we
summarize the governing equations here, and a complete description is in
our previous work~\cite{Fu2022_JFM}. The interactions satisfy a
system of partial differential equations (PDEs) that describe the
hydrodynamic interactions and hydrophobic interactions. Assuming
inertial terms are negligible, the solvent is governed by the Stokes
equations
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0},     && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity, $p$ is the pressure, $\uu_\infty$ is the
background flow, and $\mu$ is the constant viscosity. The domain $\Omega
= \Omega(t)$ is the fluid region surrounding the particles and changes
shape as the particles move. Since the particles are rigid, the solvent
velocity satisfies the no-slip rigid body motion boundary condition
\begin{align}
  \vv_i + \omega_i (\xx - \aa_i)^\perp = \uu(\xx), 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity and $\omega_i$ is the
angular velocity. Once these velocities are determined, we use
second-order Adams-Bashforth to update the particle configuration.

>>>>

%Title of paper
\title{Effects of Tunable Hydrophobicity on the Collective Hydrodynamics of Janus Particles under Flows}

@@

\begin{abstract}
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in biology. To understand how
  hydrodynamic flow may be used to actively control self-assembly of
  Janus particles (JPs), we use a model recently developed for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  background flow (JFM, 941, 2022). We investigate how various
  morphologies arise from tuning the hydrophobic distribution of the
  JP-solvent interface. We find JPs assembled into uni-lamella,
  multi-lamella and striated structures. To introduce dynamics, we
  include a linear shear flow and a steady Taylor-Green mixing flow, and
  measure the collective dynamics of JP particles in terms of their (a)
  free energy from the hydrophobic interactions between the JPs, (b)
  order parameter for the ordering of JPs in terms of alignment of their
  directors, and (c) strain parameter that captures the deformation in
  the assembly. We characterize the effective material properties of the
  JP structures and find that the uni-lamellar structures increases
  orientation order under shear flow, the multilamellar structure
  behaves as a shear thinning fluid, and the striated structure
  possesses a yield stress. These numerical results provide insights
  into dynamic control of non-equilibrium active biological systems with
  similar self-organization.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations: Hydrophobic Attraction Potential Mobility Problem} 
Following the formulation in \citep{Fu2022_JFM}, 
the governing equations are expressed as a system for the position and
orientation of a collection of rigid Janus particles. We first pose the
Stokes equations for the mobility problem giving the hydrodynamic
interactions for the particle suspension. The hydrophobic forces come
from solving a screened Laplace equation. Particle collisions are
avoided through a near-field, pair potential.

@@

To perform a single time step, we first solve for the HAP by
solving~\eqref{eq:SL}--\eqref{eq:SLbc}, and then compute the force and
torque~\eqref{eq:FandTdef}. Then the Stokes
equations~\eqref{eq:stokes}--\eqref{eq:stokesFarfield} are solved, and
the force-free and torque-free conditions are used to compute the rigid
body motion~\eqref{eq:rigid-vel}.
Once~\eqref{eq:stokes}--\eqref{eq:force} have been solved, the
velocities $(\vv_i, \omega_i)$ from~\eqref{eq:rigid-vel} are used to update the
particle positions and orientations.

@@

\section{Model parameters and boundary conditions} The interactions
generated by the HAP-mobility problem formulation
\eqref{eq:stokes}--\eqref{eq:force} lead to particle self-assembly for a
broad range of parameters and boundary conditions. Generally speaking,
the effective distance of the hydrophobic interaction is set by the
screening length $\rho$. Smaller values of the repulsion distance
$\rho_0$ decrease the distance where attraction and repulsion are in
balance and lead to more compact particle assemblies. The rate of
self-assembly is proportional to the interfacial tension $\gamma$, and
inversely proportional to solvent viscosity $\mu$; it is roughly
inversely proportional to screening length $\rho$ \cite{Fu20}.

@@

We set $\max g>0$ to guarantee at least partial hydrophobicity of the JP
surface. Then, the boundary condition (BC) can be classified into three
categories: $\min g(\xx) = 0$, $\min g(\xx) >0$, and $\min g(\xx) <0$
where the extrema are taken for $\xx \in \Gamma_i$. These three
categories correspond to three characteristic values of $b$ in
\eqref{eq:bc-type}: $b=1$, $b=2$ and $b=0$, respectively. 
%
To see the effect of the shift parameter $b$, we simulate the dynamics
of 198 particles and 60 particles in a quiescent flow as shown in
Figure~\ref{fig:relax}. Three distinct configurations emerge: (i)
bilayer (amphiphilic) for $b = 1$, (ii) multilamellar (biased hydrophobic) for $b = 2$, and (iii) striated (bipolar)
for $b = 0$. Supplementary Movie S1 shows the self-assembly process for
the three cases. In the followings the terms ``multilamellar" and ``biased hydrophobic" are used  interchangeably to refer to BC (ii), while
``striated" and ``bipolar" are used interchangeably to refer to BC (iii).

@@

Figure~\ref{fig:relax}(a) shows the initial configurations with
orientations of the 198-particle and 60-particle cases, respectively.
BC (i) with $b = 1$ mimics an amphiphilic particle. The boundary data
$g$ is everywhere nonnegative, and takes the maximum value $a$ on the
hydrophobic tail (red) the value $0$ on the hydrophilic head (green).
The interaction is attractive, and particles collectively orient their
hydrophobic sides to form bilayers (Figure~\ref{fig:relax}(b)). The
inset shows the equilibrium state for the 60-particle run. Both
$N_b = 60$ and $N_b = 198$ cases
form bilayers with multiple domains.

@@

\section{Measuring deformation}
To quantify the collective hydrodynamics of JPs,
we use the free energy $F$, a strain parameter $E$, and two scalar order
parameters $S_{2}$ and $\tilde S_2$ to measure the deformation of
particle configurations under background flows.

@@

Finally, we use the scalar order parameter $S_2$ to quantify the orientational order\cite{Selinger2016}:
\begin{equation}
  \label{eq:S2}
S_2 = \frac{1}{N_b} \sum_{i=1}^{N_b} \frac{1}{2}(3\cos^2(\theta_i - \bar \theta) - 1).
\end{equation}
Here $\bar \theta$ is a circular mean defined as the orientation of the
principal eigenvector of the matrix $\dd_1\dd_1^\top + \cdots +
\dd_{N_b}\dd_{N_b}^\top$. A value $S_2(t) = 1$ indicates that all
particle directors lie on a common axis e.g., two parallel vectors with
possibly opposite direction are ordered. A value below 1 indicates imperfect order, and zero indicates
complete disorder.

@@

%%%%%%%
%
\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{Figures/Figure5.pdf}    
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:BC1_TG} Bilayers with multiple domains in a
  Taylor-Green when $V_0 = \{0.1, 0.15, 0.2\}$ at $t=0.2\ \upmu$s. The
  pre-relaxed initial configuration is shown in the inset of panel (a).
  The streamlines are plotted in the background. Panel (d) shows the
  free energies; panel (e) shows the orientational parameter
  $\tilde{S}_2$; panel (f) shows the positional parameter $E$.}
\end{figure}
%
\subsection{Bilayer and vesicle configurations in a Taylor-Green flow}
Next, we subject the JP structures to the Taylor-Green (TG) background
flow
\begin{equation}
\label{eq:tg_flow}
\uu_\infty(\xx) = V_0\; \frac{\text{nm}}{\text{ns}}\;
\left(
-\cos\left(x/\lambda\right)
 \sin\left(y/\lambda\right)
         {\bf e}_x
         +\sin\left(x/\lambda\right)
         \cos\left(y/\lambda\right)
             {\bf e}_y\right),
\end{equation}
where $x = {\bf e}_x \cdot \xx$ and $y = {\bf e}_y \cdot \xx$
are the horizontal and vertical coordinates, respectively. 
The TG flow is a confining flow consisting of a checkerboard pattern
of cells with alternating circulation. We control the flow by the
dimensionless flow strength $V_0$ and the dimensionless cell size $l$
defining $\lambda = l$ nm. For most of the results in this subsection,
we use the parameter $l=2$.
%

@@

There is no appreciable pattern in the deformation for the
case of bilayers with multiple domains in TG flow
(Supplementary Movie S3, right panel).
Figure~\ref{fig:BC1_TG}(a)--(c)
are the snapshots when $V_0=0.1,0.15,0.2$ at $t = 0.2$ $\upmu$s. At all
flow strengths, the bilayer components are separated and mixed by the
flow in neighboring cells. In the $V_0 = 0.2$ case, the flow breaks off
components consisting of only a few particles, leading an increase in
energy (Figure~\ref{fig:BC1_TG}(d), long dashed curve). There is no
apparente structure to the change in the orientational order. There is an initial
increase in the strain parameter due to component mixing, but the
strains remains bounded since the particles stay confined to the
neighbors of the central cell.

@@

\citet{kohl-cor-che-vee22} have simulated JP suspensions in three
dimensions using the same free energy formulation of hydrophobic
attraction used in the present work. They observed spontaneous
aggregation into micelles generically and formation of bilayers when the
initial configuration was chosen sufficiently close to the final
configuration. They also studied self-assembly of bipolar electric JP,
which corresponds to our BC (iii) JP (but electric potential, instead of a water order parameter),
to define the interaction.  Due to the differing interaction,
bipolar electric particles form parallel chains that
repel. In contrast, the hydrophobic interaction between stria is
attractive. The existence of 'X'-shaped local arrangements as a local
equilibrium (Figure~\ref{fig:relax}(d), top, middle rectangle) further
suggests there may be greater diversity in the set of possible
equilibria in three-dimensions when employing hydrophobic attraction.

@@

Under a relatively weak linear shear flow, the amphiphilic JPs behave as
a unilamellar vesicle that elongates and tank-treads, with the scalar
order parameter increasing over time. The assembly of multi-lamella and
striated JP structures resemble a rigid body motion with minimal
deformation. The effective viscosity of the material quantitatively
validates this result. High shear-rate cases provide a range of critical
shear rates where the structures break apart and undergo topological
changes. We further show that free energy, scalar order parameter, and
strain are effective measures to quantitatively capture the collective hydrodynamics of
JPs under a linear shear flow.

Under a Taylor-Green flow, the amphiphilic JPs are the most interesting,
exhibiting different vesicle shapes depending on the ratio of the size
of the TG flow to the vesicle size. These results show that the shape of
the vesicle, whether square or polygonal, can be controlled by adjusting
the size of the cell in the TG flow. On the other hand, the assembly of
multi-lamella JPs and the striated JPs behave more like a rigid body
with connected subdomains, and the number of subdomains increases with
increasing strength of the TG flow. Overall, the multilamellar (BC (ii))
JP assembly behaves as a shear-thinning fluid, while the striated (BC
(iii)) JP assembly  possesses a yield stress.

@@

%%%
\subsection{Hydrophobic Attraction Potential Mobility Problem} 
The mathematical formulation is a nonlinear system for the dynamics of a
collection of rigid Janus particles. For the sake of completeness, we
summarize the governing equations here, and a complete description is in
our previous work~\cite{Fu2022_JFM}. The interactions satisfy a
system of partial differential equations (PDEs) that describe the
hydrodynamic interactions and hydrophobic interactions. Assuming
inertial terms are negligible, the solvent is governed by the Stokes
equations
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0},     && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity, $p$ is the pressure, $\uu_\infty$ is the
background flow, and $\mu$ is the (constant) solvent viscosity. The domain $\Omega
= \Omega(t)$ is the fluid region surrounding the particles and changes
shape as the particles move. Since the particles are rigid, the solvent
velocity satisfies the no-slip rigid body motion boundary condition
\begin{align}
  \vv_i + \omega_i (\xx - \aa_i)^\perp = \uu(\xx), 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity and $\omega_i$ is the
angular velocity. Once these velocities are determined, we use
second-order Adams-Bashforth to update the particle configuration.
